Comparison of Catalogs

### 2024-07-19 compare/compare.qmd
added weights

1.  [$R^2$]{.underline}: Measures the proportion of variance explained by the linear model.
2.  [Slope of the Fitted Line]{.underline}: Should be close to 1 for a 1-1 correlation.
3.  [Pearson Correlation $\rho$]{.underline}: Measures the strength and direction of the linear relationship between two variables, ranging from -1 to 1.

In simple linear regression, $R^2$ is the square of the Pearson correlation coefficient $r$.

```{python}
def compare_data(x, y):
    """
    Performs a linear comparison between two datasets.

@@

    Parameters:
    - x (array-like): The independent variable data.
    - y (array-like): The dependent variable data.

    Returns:
    tuple: A tuple containing the following elements:
        - slope (float): The slope of the fitted linear model.
        - intercept (float): The intercept of the fitted linear model.
        - r2 (float): The R-squared value, indicating the proportion of variance explained by the linear model.
        - corr (float): The Pearson correlation coefficient, measuring the linear correlation between x and y.
    """
    x_data = np.ma.array(x.value, mask=np.isnan(x.value))
    y_data =  np.ma.array(y.value, mask=np.isnan(y.value))
    # initialize a linear fitter
    fit = fitting.LinearLSQFitter()

@@

    # fit the data with the fitter
    fitted_line = fit(line_init, x_data, y_data)
    slope = fitted_line.slope.value
    intercept = fitted_line.intercept.value

@@

    Parameters:
    - x (array-like with units): Independent variable data.
    - y (array-like with units): Dependent variable data.
    xerr (array-like, optional): Error in the independent variable data. Default is 0.
    yerr (array-like, optional): Error in the dependent variable data. Default is 0.

@@

    # Handle default values for xerr and yerr
    if xerr is None:
        xerr_d = np.ma.zeros(len(x_data))
    else:
        xerr_d = np.ma.array(xerr.value, mask=np.isnan(xerr))
    if yerr is None:
        yerr_d = np.ma.zeros(len(y_data))
    else:
        yerr_d = np.ma.array(yerr.value, mask=np.isnan(yerr))

    # Perform linear comparison between the datasets
    c, m, r2, corr = compare_data(x, y)

@@

|       LCV        |        HECATE         |         Description          |        Linear Correlation         |
|:----------------:|:------------------:|:----------------:|:----------------:|
|       RVel       | V/V_VIR (with errors) | Heliocentric radial velocity | `{python} rvel_corr, rvel_corr_v` |
|       VLG        | V/V_VIR (with errors) |       Radial velocity        |  `{python} vlg_corr, vlg_corr_v`  |
| cz (with errors) | V/V_VIR (with errors) |    Heliocentric velocity     |   `{python} cz_corr, cz_corr_v`   |

@@

|      LCV      |        HECATE        |                        Description                        | Pearson Correlation \[-1,1\] |
|:---------------:|:---------------:|:---------------------:|:---------------:|
|     TType     |   T (with errors)    | Numerical Hubble type following the de Vaucouleurs system |    `{python} ttype_corr`     |
|      inc      |          I           |                     Inclination (deg)                     |     `{python} inc_corr`      |
| a26_1 (Major) | R1 (Semi-major axis) |                 angular diameter (arcmin)                 |              0               |

>>>>

1.  [$R^2$]{.underline}: Measures the proportion of variance explained by the linear model.
2.  [Slope of the Fitted Line]{.underline}: Should be close to 1 for a 1-1 correlation.[^1]
3.  [Pearson Correlation $\rho$]{.underline}: Measures the strength and direction of the linear relationship between two variables, ranging from -1 to 1. [^2]

[^1]: Fitting can be done using the uncertainties as weights. To get the standard weighting of 1/unc^2, assuming Gaussian errors, the weights to pass to the fitting are 1/unc.
[^2]: In simple linear regression, $R^2$ is the square of the Pearson correlation coefficient $\rho$.

```{python}
def compare_data(x, y, unc = None):
    """
    Performs a linear comparison between two datasets.

@@

    Parameters:
    - x (array-like): The independent variable data.
    - y (array-like): The dependent variable data.
    - unc (array-like with units, optional): The uncertainties associated with the independent variable data. Default is None.

    Returns:
    tuple: A tuple containing the following elements:
        - slope (float): The slope of the fitted linear model.
        - intercept (float): The intercept of the fitted linear model.
        - r2 (float): The R-squared value, indicating the proportion of variance explained by the linear model.
        - corr (float): The Pearson correlation coefficient, measuring the linear correlation between x and y.
    """
    x_data = np.ma.array(x.value, mask=np.isnan(x.value))
    y_data =  np.ma.array(y.value, mask=np.isnan(y.value))
    if unc is None:
        w = None
    else:
        unc = np.ma.array(unc.value, mask=np.isnan(x.value))
        w = 1/unc 
    # initialize a linear fitter
    fit = fitting.LinearLSQFitter()

@@

    # fit the data with the fitter
    fitted_line = fit(line_init, x_data, y_data, weights = w)
    slope = fitted_line.slope.value
    intercept = fitted_line.intercept.value

@@

    Parameters:
    - x (array-like with units): Independent variable data.
    - y (array-like with units): Dependent variable data.
    - xerr (array-like, optional): Error in the independent variable data. Default is 0.
    - yerr (array-like, optional): Error in the dependent variable data. Default is 0.

@@

    # Handle default values for xerr and yerr
    if xerr is None:
        xerr_d = 0
    else:
        xerr_d = np.ma.array(xerr.value, mask=np.isnan(xerr))
    if yerr is None:
        yerr_d = 0
    else:
        yerr_d = np.ma.array(yerr.value, mask=np.isnan(yerr))

    # Perform linear comparison between the datasets
    if yerr is not None:
        unc = yerr
    elif yerr is None and xerr is not None:
        unc = xerr
    else:
        unc = None
    c, m, r2, corr = compare_data(x, y, unc)

@@

|       LCV        |        HECATE         |         Description          |        Linear Correlation         |
|:----------------:|:-----------------:|:----------------:|:----------------:|
|       RVel       | V/V_VIR (with errors) | Heliocentric radial velocity | `{python} rvel_corr, rvel_corr_v` |
|       VLG        | V/V_VIR (with errors) |       Radial velocity        |  `{python} vlg_corr, vlg_corr_v`  |
| cz (with errors) | V/V_VIR (with errors) |    Heliocentric velocity     |   `{python} cz_corr, cz_corr_v`   |

@@

|      LCV      |        HECATE        |                        Description                        | Pearson Correlation \[-1,1\] |
|:---------------:|:---------------:|:-------------------:|:---------------:|
|     TType     |   T (with errors)    | Numerical Hubble type following the de Vaucouleurs system |    `{python} ttype_corr`     |
|      inc      |          I           |                     Inclination (deg)                     |     `{python} inc_corr`      |
| a26_1 (Major) | R1 (Semi-major axis) |                 angular diameter (arcmin)                 |              0               |
